---
title: "Credit"
output:
  html_document:
    theme: cerulean
    highlight: tango
    code_folding: hide
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
source('descriptive_analytics_utils.R')

## Getting the Data

# load in the data and attach the data frame
credit.df <- read.csv("german_credit_dataset.csv", header = TRUE, sep = ",")

# class should be data.frame
class(credit.df)

# get a quick peek at the data
head(credit.df)

# get dataset detailed info
str(credit.df)


```


## Data pre-processing

```{r}
# check if data frame contains NA values
sum(is.na(credit.df))

# check if total records reduced after removing rows with NA values
sum(complete.cases(credit.df))

# select variables for data transformation
categorical.vars <- c('credit.rating', 'account.balance', 'previous.credit.payment.status',
                      'credit.purpose', 'savings', 'employment.duration', 'installment.rate',
                      'marital.status', 'guarantor', 'residence.duration', 'current.assets',
                      'other.credits', 'apartment.type', 'bank.credits', 'occupation',
                      'dependents', 'telephone', 'foreign.worker')

# transform data types
credit.df <- to.factors(df = credit.df, variables = categorical.vars)

# verify transformation in data frame details
str(credit.df)

```




## Data analysis

```{r}
#load dependencies
library(car)
# access dataset features directly
attach(credit.df)


# credit.rating stats
get.categorical.variable.stats(credit.rating)
```

```{r}
# credit.rating visualizations
visualize.barchart(credit.rating)

```


```{r}
# account.balance stats and bar chart
get.categorical.variable.stats(account.balance)
visualize.barchart(account.balance)

```

```{r}
# recode classes and update data frame
new.account.balance <- recode(account.balance,
                          "1=1;2=2;3=3;4=3")
credit.df$account.balance <- new.account.balance

# contingency table and mosaic plot
get.contingency.table(credit.rating, new.account.balance)
visualize.contingency.table(credit.rating, new.account.balance)

```

```{r}
# credit.duration.months analysis
get.numeric.variable.stats(credit.duration.months)

```

```{r}
# histogram\density plot
visualize.distribution(credit.duration.months)

```

```{r}
# box plot
visualize.boxplot(credit.duration.months, credit.rating)

```

```{r}
# previous.credit.payment.status stats and bar chart
get.categorical.variable.stats(previous.credit.payment.status)
visualize.barchart(previous.credit.payment.status)

```


```{r}
# recode classes and update data frame
new.previous.credit.payment.status <- recode(previous.credit.payment.status,
                                             "0=1;1=1;2=2;3=3;4=3")
credit.df$previous.credit.payment.status <- new.previous.credit.payment.status

# contingency table
get.contingency.table(credit.rating, new.previous.credit.payment.status)

```

```{r}
# credit.purpose stats and bar chart
get.categorical.variable.stats(credit.purpose)
visualize.barchart(credit.purpose)

```


```{r}
# recode classes and update data frame
new.credit.purpose <- recode(credit.purpose,"0=4;1=1;2=2;3=3;
                                             4=3;5=3;6=3;7=4;
                                             8=4;9=4;10=4")
credit.df$credit.purpose <- new.credit.purpose

# contingency table
get.contingency.table(credit.rating, new.credit.purpose)


# credit.amount analysis
get.numeric.variable.stats(credit.amount)

# histogram\density plot
visualize.distribution(credit.amount)

```

```{r}
# box plot
visualize.boxplot(credit.amount, credit.rating)

```

```{r}


# feature: savings - recode classes and update data frame
new.savings <- recode(savings,"1=1;2=2;3=3;
                               4=3;5=4")
credit.df$savings <- new.savings

# contingency table
get.contingency.table(credit.rating, new.savings)


# feature: employment.duration - recode classes and update data frame
new.employment.duration <- recode(employment.duration,
                                  "1=1;2=1;3=2;4=3;5=4")
credit.df$employment.duration <- new.employment.duration

# contingency table
get.contingency.table(credit.rating, new.employment.duration)


# feature: installment.rate - contingency table and statistical tests
get.contingency.table(credit.rating, installment.rate,
                     stat.tests=TRUE)


# feature: marital.status - recode classes and update data frame
new.marital.status <- recode(marital.status, "1=1;2=1;3=3;4=4")
credit.df$marital.status <- new.marital.status

# contingency table
get.contingency.table(credit.rating, new.marital.status)


# feature: guarantor - recode classes and update data frame
new.guarantor <- recode(guarantor, "1=1;2=2;3=2")
credit.df$guarantor <- new.guarantor

# perform statistical tests
fisher.test(credit.rating, new.guarantor)
chisq.test(credit.rating, new.guarantor)


# perform statistical tests for residence.duration
fisher.test(credit.rating, residence.duration)
chisq.test(credit.rating, residence.duration)


# perform statistical tests for current.assets
fisher.test(credit.rating, current.assets)
chisq.test(credit.rating, current.assets)


# age analysis
get.numeric.variable.stats(age)

# histogram\density plot
visualize.distribution(age)

# box plot
visualize.boxplot(age, credit.rating)
```


```{r}

# feature: other.credits - recode classes and update data frame
new.other.credits <- recode(other.credits, "1=1;2=1;3=2")
credit.df$other.credits <- new.other.credits

# perform statistical tests
fisher.test(credit.rating, new.other.credits)
chisq.test(credit.rating, new.other.credits)


# perform statistical tests for apartment.type
fisher.test(credit.rating, apartment.type)
chisq.test(credit.rating, apartment.type)


# feature: bank.credits - recode classes and update data frame
new.bank.credits <- recode(bank.credits, "1=1;2=2;3=2;4=2")
credit.df$bank.credits <- new.bank.credits

# perform statistical tests
fisher.test(credit.rating, new.bank.credits)
chisq.test(credit.rating, new.bank.credits)


# perform statistical tests for occupation
fisher.test(credit.rating, occupation)
chisq.test(credit.rating, occupation)


# perform statistical tests for dependents
fisher.test(credit.rating, dependents)
chisq.test(credit.rating, dependents)


# perform statistical tests for telephone
fisher.test(credit.rating, telephone)
chisq.test(credit.rating, telephone)


# perform statistical tests for foreign.worker
fisher.test(credit.rating, foreign.worker)
chisq.test(credit.rating, foreign.worker)

```



```{r}

## Save the transformed dataset
write.csv(file='credit_dataset_final1.csv', x = credit.df,
          row.names = F)
```


# Credit Risk Detection and Prediction - Predictive Analytics (Chapter 6)

## Data preparation
```{r}
# load the dataset into data frame
credit.df <- read.csv("credit_dataset_final.csv", header = TRUE, sep = ",")

## data type transformations - factoring
to.factors <- function(df, variables){
  for (variable in variables){
    df[[variable]] <- as.factor(df[[variable]])
  }
  return(df)
}

## normalizing - scaling
scale.features <- function(df, variables){
  for (variable in variables){
    df[[variable]] <- scale(df[[variable]], center=T, scale=T)
  }
  return(df)
}

# normalize variables
numeric.vars <- c("credit.duration.months", "age", "credit.amount")
credit.df <- scale.features(credit.df, numeric.vars)
# factor variables
categorical.vars <- c('credit.rating', 'account.balance', 'previous.credit.payment.status',
                      'credit.purpose', 'savings', 'employment.duration', 'installment.rate',
                      'marital.status', 'guarantor', 'residence.duration', 'current.assets',
                      'other.credits', 'apartment.type', 'bank.credits', 'occupation', 
                      'dependents', 'telephone', 'foreign.worker')
credit.df <- to.factors(df=credit.df, variables=categorical.vars)

# split data into training and test datasets in 60:40 ratio
indexes <- sample(1:nrow(credit.df), size=0.6*nrow(credit.df))
train.data <- credit.df[indexes,]
test.data <- credit.df[-indexes,]
```

## Feature Selection

```{r}
library(caret)  # feature selection algorithm
library(randomForest) # random forest algorithm

# rfe based feature selection algorithm
run.feature.selection <- function(num.iters=20, feature.vars, class.var){
  set.seed(10)
  variable.sizes <- 1:10
  control <- rfeControl(functions = rfFuncs, method = "cv", 
                        verbose = FALSE, returnResamp = "all", 
                        number = num.iters)
  results.rfe <- rfe(x = feature.vars, y = class.var, 
             sizes = variable.sizes, 
             rfeControl = control)
  return(results.rfe)
}

# run feature selection
rfe.results <- run.feature.selection(feature.vars=train.data[,-1], 
                                     class.var=train.data[,1])
# view results
rfe.results
```
```{r}
varImp(rfe.results)
```

#Models

## Logistic Regression

For our model, since we
are dealing with a binary classification problem, we will be dealing with binomial
logistic regression.

```{r}
library(caret) # model training and evaluation
library(ROCR) # model evaluation
source("performance_plot_utils.R") # plotting metric results

## separate feature and class variables
test.feature.vars <- test.data[,-1]
test.class.var <- test.data[,1]

# build a logistic regression model
formula.init <- "credit.rating ~ ."
formula.init <- as.formula(formula.init)
lr.model <- glm(formula=formula.init, data=train.data, family="binomial")

# view model details
summary(lr.model)

# perform and evaluate predictions
lr.predictions <- predict(lr.model, test.data, type="response")
lr.predictions <- round(lr.predictions)
confusionMatrix(data=lr.predictions, reference=test.class.var, positive='1')


## glm specific feature selection
formula <- "credit.rating ~ ."
formula <- as.formula(formula)
control <- trainControl(method="repeatedcv", number=10, repeats=2)
model <- train(formula, data=train.data, method="glm", 
               trControl=control)
importance <- varImp(model, scale=FALSE)
plot(importance)


# build new model with selected features
formula.new <- "credit.rating ~ account.balance + credit.purpose + previous.credit.payment.status 
                                + savings + credit.duration.months"
formula.new <- as.formula(formula.new)
lr.model.new <- glm(formula=formula.new, data=train.data, family="binomial")

# view model details
summary(lr.model.new)

# perform and evaluate predictions 
lr.predictions.new <- predict(lr.model.new, test.data, type="response") 
lr.predictions.new <- round(lr.predictions.new)
confusionMatrix(data=lr.predictions.new, reference=test.class.var, positive='1')



## model performance evaluations

# plot best model evaluation metric curves
lr.model.best <- lr.model
lr.prediction.values <- predict(lr.model.best, test.feature.vars, type="response")
predictions <- prediction(lr.prediction.values, test.class.var)
par(mfrow=c(1,2))
plot.roc.curve(predictions, title.text="LR ROC Curve")
plot.pr.curve(predictions, title.text="LR Precision/Recall Curve")
```

## SVM Classifier

```{r}
library(e1071) # svm model
library(caret) # model training\optimizations
library(kernlab) # svm model for hyperparameters
library(ROCR) # model evaluation
source("performance_plot_utils.R") # plot model metrics


## separate feature and class variables
test.feature.vars <- test.data[,-1]
test.class.var <- test.data[,1]

## build initial model with training data
formula.init <- "credit.rating ~ ."
formula.init <- as.formula(formula.init)
svm.model <- svm(formula=formula.init, data=train.data, 
                 kernel="radial", cost=100, gamma=1)

## view inital model details
summary(svm.model)

## predict and evaluate results
svm.predictions <- predict(svm.model, test.feature.vars)
confusionMatrix(data=svm.predictions, reference=test.class.var, positive="1")


## svm specific feature selection
formula.init <- "credit.rating ~ ."
formula.init <- as.formula(formula.init)
control <- trainControl(method="repeatedcv", number=10, repeats=2)
model <- train(formula.init, data=train.data, method="svmRadial", 
               trControl=control)
importance <- varImp(model, scale=FALSE)
plot(importance, cex.lab=0.5)


## build new model with selected features
formula.new <- "credit.rating ~ account.balance + credit.duration.months +
                            savings + previous.credit.payment.status +
                            credit.amount"
formula.new <- as.formula(formula.new)
svm.model.new <- svm(formula=formula.new, data=train.data, 
                 kernel="radial", cost=10, gamma=0.25)

## predict results with new model on test data
svm.predictions.new <- predict(svm.model.new, test.feature.vars)

## new model performance evaluation
confusionMatrix(data=svm.predictions.new, reference=test.class.var, positive="1")



## hyperparameter optimizations

# run grid search
cost.weights <- c(0.1, 10, 100)
gamma.weights <- c(0.01, 0.25, 0.5, 1)
tuning.results <- tune(svm, formula.new, 
                       data = train.data, kernel="radial", 
                       ranges=list(cost=cost.weights, gamma=gamma.weights))

# view optimization results
print(tuning.results)

# plot results
plot(tuning.results, cex.main=0.6, cex.lab=0.8,xaxs="i", yaxs="i")

# get best model and evaluate predictions
svm.model.best = tuning.results$best.model
svm.predictions.best <- predict(svm.model.best, test.feature.vars)
confusionMatrix(data=svm.predictions.best, reference=test.class.var, positive="1")


# plot best model evaluation metric curves
svm.predictions.best <- predict(svm.model.best, test.feature.vars, decision.values = T)
svm.prediction.values <- attributes(svm.predictions.best)$decision.values
predictions <- prediction(svm.prediction.values, test.class.var)
par(mfrow=c(1,2))
plot.roc.curve(predictions, title.text="SVM ROC Curve")
plot.pr.curve(predictions, title.text="SVM Precision/Recall Curve")



## model optimizations based on ROC

# data transformation
transformed.train <- train.data
transformed.test <- test.data
for (variable in categorical.vars){
  new.train.var <- make.names(train.data[[variable]])
  transformed.train[[variable]] <- new.train.var
  new.test.var <- make.names(test.data[[variable]])
  transformed.test[[variable]] <- new.test.var
}
transformed.train <- to.factors(df=transformed.train, variables=categorical.vars)
transformed.test <- to.factors(df=transformed.test, variables=categorical.vars)
transformed.test.feature.vars <- transformed.test[,-1]
transformed.test.class.var <- transformed.test[,1]

# view data to understand transformations
summary(transformed.train$credit.rating)

# build optimal model based on AUC
grid <- expand.grid(C=c(1,10,100), 
                    sigma=c(0.01, 0.05, 0.1, 0.5, 1))
ctr <- trainControl(method='cv', number=10,
                    classProbs=TRUE,
                    summaryFunction=twoClassSummary)
svm.roc.model <- train(formula.init, transformed.train,
                       method='svmRadial', trControl=ctr, 
                       tuneGrid=grid, metric="ROC")

# predict and evaluate model performance
predictions <- predict(svm.roc.model, transformed.test.feature.vars)
confusionMatrix(predictions, transformed.test.class.var, positive = "X1")


## plot model evaluation metric curves
svm.predictions <- predict(svm.roc.model, transformed.test.feature.vars, type="prob")
svm.prediction.values <- svm.predictions[,2]
predictions <- prediction(svm.prediction.values, test.class.var)
par(mfrow=c(1,2))
plot.roc.curve(predictions, title.text="SVM ROC Curve")
plot.pr.curve(predictions, title.text="SVM Precision/Recall Curve")
```

It is quite pleasing to see that the AUC has indeed increased from 0.69 earlier to
0.74 now, which means the AUC based optimization algorithm indeed worked,
since it has given better performance than the previous model in all the aspects we
have been tracking.

## Decision trees